---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(lubridate)

rawdata<-read.csv("activity.csv")
rawdata$date<-as.Date(rawdata$date)
```


## What is mean total number of steps taken per day?
```{r total_steps}
naomit_data<-na.omit(rawdata)
steps_sum<-tapply(naomit_data$steps, naomit_data$date, sum)
count_table<-data.frame(Date = names(steps_sum), Steps = steps_sum)
count_table$Date<-as.Date(count_table$Date)

hist(count_table$Steps, xlab = "Total Steps Taken per Day", main = "Histogram of Steps Taken per Day")
```


### Mean steps per day
```{r mean_steps}
mean_steps<-mean(count_table$Steps)
print(mean_steps)
```

### Median steps per day
```{r median_steps}
median_steps<-median(count_table$Steps)
print(median_steps)
```

## What is the average daily activity pattern?
```{r avg_daily_patterm}
avg_steps<-tapply(naomit_data$steps, naomit_data$interval, mean)
avgsteps_table<-data.frame(Interval = names(avg_steps), Average = avg_steps)

plot(avgsteps_table$Interval, avgsteps_table$Average, type = "l", xlab = "Interval", ylab = "Average Steps Taken", main = "Average Number of Steps Taken per 5 min Interval (Oct - Nov 2012)")
```

### Interval with maximum steps
```{r max_interval}
maxsteps_interval<-subset(avgsteps_table, avgsteps_table$Average == max(avgsteps_table$Average))[1,1]
print(maxsteps_interval)
```

### Number of missing values
```{r num_missing}
num_na<-sum(is.na(rawdata$steps))
print(num_na)
```

### Imputing missing values
Use mean for that interval if not available use mean for that day, and if day mean not available use zero
```{r imputation}
rawdata_fixed<-rawdata
for(i in 1:17568){
  if(is.na(rawdata_fixed[i,1])){
    value<-subset(avgsteps_table$Average, avgsteps_table$Interval == rawdata_fixed[i,3])

    if(!is.null(value)){
      rawdata_fixed[i,1]<-value
    }
    else{
      datetemp<-rawdata_fixed[i,2]
      subset_temp<-subset(count_table$Steps, count_table$Date == datetemp)
      if(!is.null(subset_temp) & length(subset_temp) != 0){
        rawdata_fixed[i,1]<-subset_temp
      }
      else{
        rawdata_fixed[i,1]<-0
      }
    } 
  }
}
```

### Number of steps taken per day with imputed data added
```{r total_steps2}
steps_sum2<-tapply(rawdata_fixed$steps, rawdata_fixed$date, sum)
count_table2<-data.frame(Date = names(steps_sum2), Steps = steps_sum2)
count_table2$Date<-as.Date(count_table2$Date)
hist(count_table2$Steps, xlab = "Total Steps Taken per Day", main = "Histogram of Steps Taken per Day (with Imputed Data)")
```

### Mean steps per day
```{r mean_steps2}
mean_steps2<-mean(count_table2$Steps)
print(mean_steps2)
```

### Median steps per day
```{r median_steps2}
median_steps2<-median(count_table2$Steps)
print(median_steps2)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r wkday_vs_wkend}
rawdata_fixed$Weekday<-weekdays(rawdata_fixed$date, abbreviate = TRUE)
rawdata_fixed$Is_weekend<-(rawdata_fixed$Weekday == "Sat" | rawdata_fixed$Weekday == "Sun")

sub_wkday<-subset(rawdata_fixed, rawdata_fixed$Is_weekend == FALSE)
sub_wkend<-subset(rawdata_fixed, rawdata_fixed$Is_weekend == TRUE)

avg_steps_wkday<-tapply(sub_wkday$steps, sub_wkday$interval, mean)
avgsteps_table_wkday<-data.frame(Interval = names(avg_steps_wkday), Average = avg_steps_wkday, Is_weekend = "Weekdays")

avg_steps_wkend<-tapply(sub_wkend$steps, sub_wkend$interval, mean)
avgsteps_table_wkend<-data.frame(Interval = names(avg_steps_wkend), Average = avg_steps_wkend, Is_weekend = "Weekends")

finalweekdays_table<-rbind(avgsteps_table_wkday, avgsteps_table_wkend)
finalweekdays_table$Interval<-as.numeric(finalweekdays_table$Interval)

qplot(Interval, Average, data = finalweekdays_table, geom = "line", facets = .~Is_weekend, ylab = "Average Steps Taken", main = "Average Steps Taken per 5 min Interval - Weekdays vs Weekends")
```

End of Document